---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
require(ggplot2)
library(scales)
library(dplyr)
setwd("/Users/alibagherpour/Documents/LearnR/Reproducible Research/RepData_PeerAssessment1")
```

## Loading and preprocessing the data
```{r}
ActivityData <- read.csv("./activity.csv")
```



## What is mean total number of steps taken per day?

```{r}
StepPerDay <- aggregate(ActivityData$steps, by=list(ActivityData$date), FUN = sum,
                        na.rm=FALSE)
qplot(x, data=StepPerDay, geom="histogram", ylab = "Counts",
      xlab = "Number of steps per day")
```

Mean and median of the total number of steps taken per day
```{r}
MeanSteps <- mean(StepPerDay$x,na.rm=TRUE)
MedianSteps <- median(StepPerDay$x,na.rm=TRUE)
```
Mean and median of the total number of steps taken per day are `r format(MeanSteps,scientific=FALSE)` 
and `r MedianSteps`, respectively.



## What is the average daily activity pattern?
```{r}
ActivityData$interval <- formatC(ActivityData$interval, width = 4, format = "d", flag = "0")
ActivityData$interval_initial <- ActivityData$interval
ActivityData$interval_initial <- 
        as.POSIXct(paste0(ActivityData$date,ActivityData$interval,sep=" "), 
                   format="%Y-%m-%d %H%M")
ActivityData$interval <- 
        as.POSIXct(ActivityData$interval, format="%H%M")

StepsTimeSeries <- aggregate(ActivityData$steps, by=list(ActivityData$interval), 
                             FUN = "mean",na.rm=TRUE)

ggplot(StepsTimeSeries, aes(Group.1,x)) + geom_line() + 
        scale_x_datetime(labels = date_format("%H:%M")) + xlab("Time") +
        ylab("Average number of steps") 
MaxActivityTime <- StepsTimeSeries$Group.1[which.max(StepsTimeSeries$x)]
```
Maximum number of steps on average happens at `r format(MaxActivityTime,"%H:%M")` AM.

## Imputing missing values
The total number of missing values in dataset is `r sum(is.na(ActivityData$steps))`. The total number of reported values is `r length(ActivityData$steps)`.

```{r}
ActivityDataReplaceNA <- ActivityData %>% 
            group_by(strftime(ActivityData$interval, format="%H:%M:%S")) %>%
            mutate(steps= replace(steps, is.na(steps), mean(steps, na.rm=TRUE)))
```
As shown in the code, the mean for 5-minute intervals was used to fill in all of the missing values in the dataset.


```{r}
StepPerDayReplaceNA <- aggregate(ActivityDataReplaceNA$steps, 
                                 by=list(ActivityDataReplaceNA$date),
                                 FUN = sum,na.rm=FALSE)
qplot(x, data=StepPerDayReplaceNA, geom="histogram", ylab = "Counts",
      xlab = "Number of steps per day")
```

Calculating mean and median of the total number of steps taken per day when the missing data was replaced:
```{r}
MeanStepsReplaceNA <- mean(StepPerDayReplaceNA$x,na.rm=FALSE)
MedianStepsReplaceNA <- median(StepPerDayReplaceNA$x,na.rm=FALSE)
```
Mean and median of the total number of steps taken per day are `r format(MeanStepsReplaceNA,scientific=FALSE)` and `r format(MedianStepsReplaceNA,scientific=FALSE)`, respectively.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
ActivityDataReplaceNA$wDay <- c('weekend', 'weekday')[(weekdays(ActivityDataReplaceNA$interval_initial) %in% weekdays1)+1L]
weekday_averaged <- 
        summarise(group_by(ActivityDataReplaceNA, interval, wDay),mean=mean(steps))
h <- ggplot(weekday_averaged, aes(interval, mean)) + geom_line() + facet_grid(. ~ wDay)
h + scale_x_datetime(labels = date_format("%H:%M")) + xlab("Time") +
        ylab("Average number of steps") 
```





